---
title: "Quantifying the Impact of Residential Solar on ASEAN CO₂ Emissions"
output: html_document
date: "2025-07-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Load necessary libraries
library(tidyverse) # Use for data science 
library(scales) # Use for improving formats/axis
library(ggrepel) # Use for fixing overlapping

### **Introduction**
This report investigates the potential for residential rooftop solar panels to reduce CO₂ emissions from the electricity sector across the Association of Southeast Asian Nations (ASEAN). By integrating datasets on carbon emissions, grid intensity, and solar energy potential, we quantify the abatement capacity of a standard 5 kW solar PV system and estimate the scale of adoption required to achieve a meaningful 5% reduction in power-sector emissions.

```{r load-data}
# Load the pre-processed datasets
co2_emissions <- read_csv("../data-raw/interim/asean_co2_emissions.csv")
grid_emission <- read_csv("../data-raw/interim/asean_GridEmission.csv")
pv_potential <- read_csv("../data-raw/interim/asean_PV.csv")
summary_final <- read_csv("../data/asean_summary_final.csv")

```


```{r plot-co2-emissions}
# Prepare data for plotting
power_emissions_2023 <- co2_emissions %>%
  filter(Sector == "Power Industry", Year == 2023) %>%
  mutate(
    co2_emissions_mt = co2_emissions_g / 1e12, # Convert grams to million metric tons
    country = fct_reorder(country, co2_emissions_mt)
    )

# Create the plot
ggplot(power_emissions_2023, aes(x = co2_emissions_mt, y = country, fill = co2_emissions_mt)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste(round(co2_emissions_mt), "M")), hjust = -0.1, size = 3.5, color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkred") +
  scale_x_continuous(labels = scales::comma_format(suffix = " M tons"), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Power Industry CO2 Emissions in ASEAN (2023)",
    x = "CO2 Emissions (Million Metric Tons)",
    y = "Country"
  ) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.y = element_blank())
```
```{r Analysis on -co2-emissions}
### **1. Analysis of Carbon Emission Intensity**

The analysis of the emissions data confirms that the **"Power Industry"** is the primary driver of CO₂ emissions in most ASEAN nations, making it the critical target for decarbonization efforts. The data from 2023 reveals a significant disparity in CO₂ emissions from the power sector across the region. **Indonesia** is the region's largest emitter by a substantial margin, accounting for approximately **270 million metric tons of CO₂**. It is followed by other major contributors including **Vietnam (155M tons)**, **Malaysia (117M tons)**, **Thailand (80M tons)**, and the **Philippines (83M tons)**. These five nations are responsible for the vast majority of the region's power-sector emissions.


///////////////////////////////////////////////////////////////////////////////////////////
### After Analyzing the Grid Emission Plot
Connecting this to grid intensity, the data shows a clear correlation: countries with high overall emissions also have highly carbon-intensive grids. Indonesia, the top emitter, also has the highest emission factor, reinforcing the fact that its reliance on fossil fuels for power generation is the core issue.

```

### **2. Analysis of Grid Emission Intensity**

```{r plot-grid-emission}
# Prepare data for plotting
grid_intensity_data <- grid_emission %>%
  filter(!is.na(ef_gco2_per_kwh)) %>%
  mutate(country = fct_reorder(country, ef_gco2_per_kwh))

# Create the plot
ggplot(grid_intensity_data, aes(x = ef_gco2_per_kwh, y = country, fill = ef_gco2_per_kwh)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = round(ef_gco2_per_kwh, 1)), hjust = -0.2, size = 4) +
  scale_fill_gradient(low = "seagreen3", high = "firebrick3") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Grid Emission Intensity Across ASEAN",
    x = "Emission Factor (gCO₂ per kWh)",
    y = "Country"
  ) +
  theme_minimal(base_size = 14) +
    theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.y = element_blank())
```
```{r analysis on-grid-emission}

The grid emission factor (EF) is a critical metric that determines the effectiveness of solar energy in displacing carbon. The analysis reveals:

*   **High-Intensity Grids:** Indonesia's grid is the most carbon-intensive at **778 gCO₂/kWh**, followed by Malaysia (670 gCO₂/kWh). This is indicative of a heavy reliance on fossil fuels, particularly coal, for electricity generation. Every kWh of solar energy produced in these countries has a disproportionately high impact on reducing emissions.
*   **Low-Intensity Grids:** Conversely, Myanmar (395 gCO₂/kWh) has the least carbon-intensive grid among the countries with available data, likely due to a different energy mix that may include more hydropower.
*   **Data Gaps:** It is important to note that grid emission data was not available for Brunei, preventing a direct calculation of its solar abatement potential in this study.
```


### **3. Analysis of Solar Energy Potential**
```{r plot-pv-factor}
# Prepare data for plotting
pv_factor_data <- pv_potential %>%
  mutate(country = fct_reorder(country, solar_capacity_factor))

# Create the plot
ggplot(pv_factor_data, aes(x = solar_capacity_factor, y = country)) +
  geom_col(fill = "orange") +
  geom_text(aes(label = scales::percent(solar_capacity_factor, accuracy = 0.1)), hjust = -0.2, size = 4) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Solar PV Capacity Factor in ASEAN Countries",
    x = "Solar Capacity Factor",
    y = "Country"
  ) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.y = element_blank())
```
```{r analysis on-sfc}
The analysis of solar capacity factors—the ratio of actual energy output over a year to the theoretical maximum—reveals that ASEAN is a prime region for solar energy.

**Uniformly High Potential:** The solar capacity factors are remarkably consistent across the region, ranging from a high of **~17% in Cambodia and Thailand** to a low of **~14.3% in Vietnam**. 
This uniformity simplifies regional analysis and strongly suggests that solar PV technology is a viable and predictable energy source for all ASEAN nations. The `solar_capacity_factor` is a key input for the R script, which uses it to calculate the annual electricity generation of a standard 5 kW residential system.


```


### **4. Analysis of Household Data**
While the number of required solar installations is a key output of this analysis, its practical meaning becomes clear only when contextualized with demographic data. The asean_Household.csv dataset, which contains the total number of households per country, serves this critical purpose.

For instance, this study finds that the Philippines requires approximately 1.16 million homes to be equipped with 5 kW solar systems to meet a 5% emissions reduction target. By using the household data, a policymaker can immediately translate this absolute figure into a market penetration rate (e.g., as a percentage of total households). This context is indispensable for assessing the feasibility of such a program, designing targeted subsidy schemes, and understanding the true scale of the national effort required to achieve these climate goals. It transforms a raw number into an actionable policy metric.

### **5. Analysis of Solar Abatement Potential**

This final stage of analysis combines the previous datasets, which synthesizes emissions, grid, and solar data across multiple scenarios. For this report, we will focus on a central scenario: achieving a 5% reduction in 2023 power-sector emissions using 5 kW residential solar systems.

```{r calculate-abatement-potential}
# Create a single, comprehensive data frame by joining the sources and then calculating the required metrics.

analysis_scenario <- summary_final %>%
  filter(PV_SYSTEM_SIZE == 5, TARGET_REDUCTION == 0.05) %>%
  # Ensure data is clean for plotting
  filter(!is.na(co2_avoided_per_home) & !is.na(homes_required))
```

#### **Finding 1: A Single Home Solar System Has the Greatest CO₂ Impact in Indonesia and Malaysia**

It calculates the CO₂ avoided annually by a standard 5 kW residential solar system. Due to its high grid intensity, a single solar installation in Indonesia can prevent approximately 5.2 metric tons of CO₂ per year. Malaysia follows, with each system avoiding 4.5 metric tons annually. In contrast, the same system in Myanmar, with its cleaner grid, avoids a smaller but still substantial 2.9 tons. This highlights that solar investments yield the highest carbon returns in countries with the "dirtiest" grids.

```{r plot-co2-avoided}
co2_avoided_data <- analysis_scenario %>%
  mutate(
    co2_avoided_tons = co2_avoided_per_home / 1e6, # Convert grams to metric tons
    country = fct_reorder(country, co2_avoided_tons)
  )

# Create the plot
ggplot(co2_avoided_data, aes(x = co2_avoided_tons, y = country, fill = co2_avoided_tons)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = round(co2_avoided_tons, 1)), hjust = -0.2, size = 4) +
  scale_fill_gradient(low = "lightgreen", high = "darkgreen") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Annual CO₂ Abatement per 5 kW Residential Solar Panel",
    x = "CO₂ Avoided (Metric Tons per Home per Year)",
    y = "Country"
  ) +
  theme_minimal(base_size = 14) +
    theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.y = element_blank())
```
#### **Finding 2: Achieving a Modest 5% Emissions Reduction Requires a Massive, Coordinated Effort**

```{r plot-homes-required}
# Prepare data for plotting
homes_required_data <- analysis_scenario %>%
  mutate(country = fct_reorder(country, homes_required))

# Create the plot
ggplot(homes_required_data, aes(x = homes_required, y = country, fill = homes_required)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = scales::comma(round(homes_required, -3))), hjust = -0.1, size = 3.5) +
  scale_fill_gradient(low = "lightgoldenrod1", high = "darkorange3") +
  scale_x_continuous(labels = scales::comma_format(), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Number of Homes Needed to Achieve a 5% Reduction in Power-Sector CO₂ Emissions",
    x = "Number of Homes Required",
    y = "Country"
  ) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.y = element_blank())
```

```{r findings B}
This is the headline finding of the study. To reduce the 2023 power-sector emissions by just 5%, the number of homes that would need to adopt a 5 kW solar system is immense:

*   **Indonesia:** 2.6 million homes
*   **Vietnam:** 2.05 million homes
*   **Malaysia:** 1.3 million homes
*   **Philippines:** 1.16 million homes

For smaller emitters like Cambodia, the absolute number is lower at approximately 52,000 homes, but this still represents a major national project requiring significant investment and policy support. These figures underscore the massive scale of the challenge and the need for immediate, large-scale action.

```

### **Conclusion and Policy Recommendations**

Based on this quantitative analysis, the following policy recommendations are proposed:

1. **Adopt a Two-Pronged Prioritization Strategy:**

- The data clearly shows that not all countries offer the same return on investment. Policy and climate finance should be prioritized based on two distinct metrics:
For Maximum Impact: To achieve the greatest CO₂ reduction per solar panel installed, efforts should be concentrated in countries with the most carbon-intensive grids. As shown in our "Annual CO₂ Abatement" chart, a single 5 kW system in Indonesia (5.2 tons/year) and Malaysia (4.5 tons/year) has a disproportionately high impact.

- To Tackle the Largest Emitters: To address the sheer volume of emissions, large-scale national programs are critically needed in the region's biggest polluters. The "Homes Required" analysis 
identifies Indonesia (2.6M homes), Vietnam (2.05M homes), Malaysia (1.3M homes), and the Philippines (1.16M homes) as the nations requiring the most significant mobilization.

2. **Capitalize on Uniform Solar Potential for Regional Collaboration:**

= The "Solar PV Capacity Factor" chart demonstrates that the entire ASEAN region is a prime location for solar energy, with consistently high capacity factors ranging from 14% to 17%. This uniformity is a strategic asset. ASEAN nations should leverage this shared advantage to:
Establish a regional purchasing bloc to lower the cost of solar panels and inverters through bulk procurement.

- Develop common technical standards for residential solar installations to ensure quality, safety, and grid compatibility.
Create a platform for sharing best practices on grid integration, policy design, and installer training, accelerating the learning curve for all member states.
